--- 
title: "General Education Report^[Report number 1710, DOI [10.17605/OSF.IO/35GSR](https://doi.org/10.17605/OSF.IO/35GSR)]"
subtitle: "Quantitative Literacy Outcome QNT3, Fall 2017"
author: 
  - "Dr. Clifton Franklund"
  - "General Education Coordinator"
date: "Fall 2017"
output: 
  bookdown::gitbook:
    includes:
      after_body: disqus.html
description: This report summarizes the student learning on Quantitative Literacy FLO QNT3 for the Fall of 2017.
documentclass: article
github-repo: WeeBeasties/1710
link-citations: yes
bibliography: references.bib
site: bookdown::bookdown_site
biblio-style: apalike
---
\addtolength{\headheight}{0.7cm}
\thispagestyle{fancyplain}
\lhead{\includegraphics[height=0.5cm]{art/logo.png}}
\rhead{}
\renewcommand{\headrulewidth}{0pt}

```{R packages, echo=FALSE, message=FALSE, warning=FALSE}

#----------------------------------------------------------------------------------------
#	INSTALL PACKAGES
#----------------------------------------------------------------------------------------

library(tidyverse)      # the tidyverse
library(pander) 	# code layout
library(moments)        # calculate skew, kurtosis, etc.
library(weights)        # calculate weighted t-test
library(forestplot)     # present meta-analysis
```

```{R functions, echo=FALSE, message=FALSE, warning=FALSE}

#----------------------------------------------------------------------------------------
#	DEFINE FUNCTIONS
#----------------------------------------------------------------------------------------

# Calculate mode
myMode <- function(x) {
  ux <- unique(x)
  ux[which.max(tabulate(match(x, ux)))]
}

# Adds legends above figures (used for barplot)
add_legend <- function(...) {
  opar <- par(fig=c(0, 1, 0, 1), oma=c(0, 0, 0, 0),
    mar=c(0, 0, 0, 0), new=TRUE)
  on.exit(par(opar))
  plot(0, 0, type='n', bty='n', xaxt='n', yaxt='n')
  legend(...)
}

#  Computes the variance of a weighted mean following Cochran 1977 definition
#  Code found online at:
#  http://stats.stackexchange.com/questions/25895/computing-standard-error-in-weighted-mean-estimation
weighted.var.se <- function(x, w, na.rm=FALSE)
	{
	if (na.rm) { w <- w[i <- !is.na(x)]; x <- x[i] }
	n = length(w)
	xWbar = weighted.mean(x,w,na.rm=na.rm)
	wbar = mean(w)
	out = sqrt(n/((n-1)*sum(w)^2)*(sum((w*x-wbar*xWbar)^2)-2*xWbar*sum((w-wbar)*(w*x-wbar*xWbar))+xWbar^2*sum((w-wbar)^2)))
	low = xWbar-(out*1.96)
	high = xWbar+(out*1.96)
	myOutput <- c(mean=format(round(xWbar,2),nsmall=2),low=format(round(low,2),nsmall=2),high=format(round(high,2),nsmall=2))
	return(myOutput)
}

# Computes omega squared (effect size) for an ANOVA analysis
omega_sq <- function(aovm){
    sum_stats <- summary(aovm)[[1]]
    SSm <- sum_stats[["Sum Sq"]][1]
    SSr <- sum_stats[["Sum Sq"]][2]
    DFm <- sum_stats[["Df"]][1]
    MSr <- sum_stats[["Mean Sq"]][2]
    W2 <- (SSm-DFm*MSr)/(SSm+SSr+MSr)
    return(W2)
}


## Summarizes data.
## Gives count, mean, standard deviation, standard error of the mean, and confidence interval (default 95%).
##   data: a data frame.
##   measurevar: the name of a column that contains the variable to be summariezed
##   groupvars: a vector containing names of columns that contain grouping variables
##   na.rm: a boolean that indicates whether to ignore NA's
##   conf.interval: the percent range of the confidence interval (default is 95%)
summarySE <- function(data=NULL, measurevar, groupvars=NULL, na.rm=FALSE,
                      conf.interval=.95, .drop=TRUE) {
    library(plyr)

    # New version of length which can handle NA's: if na.rm==T, don't count them
    length2 <- function (x, na.rm=FALSE) {
        if (na.rm) sum(!is.na(x))
        else       length(x)
    }

    # This does the summary. For each group's data frame, return a vector with
    # N, mean, and sd
    datac <- ddply(data, groupvars, .drop=.drop,
      .fun = function(xx, col) {
        c(N    = length2(xx[[col]], na.rm=na.rm),
          mean = mean   (xx[[col]], na.rm=na.rm),
          sd   = sd     (xx[[col]], na.rm=na.rm)
        )
      },
      measurevar
    )

    # Rename the "mean" column    
    datac <- rename(datac, c("mean" = measurevar))

    datac$se <- datac$sd / sqrt(datac$N)  # Calculate standard error of the mean

    # Confidence interval multiplier for standard error
    # Calculate t-statistic for confidence interval: 
    # e.g., if conf.interval is .95, use .975 (above/below), and use df=N-1
    ciMult <- qt(conf.interval/2 + .5, datac$N-1)
    datac$ci <- datac$se * ciMult

    return(datac)
}
```

```{R data, echo=FALSE, message=FALSE, warning=FALSE}

#----------------------------------------------------------------------------------------
#	LOAD DATA
#----------------------------------------------------------------------------------------
reportData <- read_csv("data/mathData.csv")
reportData$Gender[is.na(reportData$Gender)] <- "NA"
reportData$QNT1 <- as.numeric(reportData$QNT1)
reportData$QNT2 <- as.numeric(reportData$QNT2)
reportData$QNT3 <- as.numeric(reportData$QNT3)
reportData$QNT4 <- as.numeric(reportData$QNT4)
reportData <- reportData %>% 
	filter(!is.na(QNT3))
reportData$Semester <- "Fall 2017"

reportDataGender <- reportData %>% filter(!is.na(Gender))
reportDataGender$Level <- as.factor(reportDataGender$Level)
reportDataGender$`Race/Ethnicity`[reportDataGender$`Race/Ethnicity`=="American Indian/Alaskan Native"] <- "Other"
reportDataGender$`Race/Ethnicity`[reportDataGender$`Race/Ethnicity`=="Unknown"] <- "Other"
reportDataGender$`Race/Ethnicity`[is.na(reportDataGender$`Race/Ethnicity`)] <- "Other"

reportDataTable <- reportData %>%
	group_by(Order) %>%
	summarise(Prefix = Prefix[1], Level = Level[1], n = length(QNT3), mean = mean(QNT3), lower = mean(QNT3)-(1.96*sd(QNT3)/sqrt(length(QNT3))), upper = mean(QNT3)+(1.96*sd(QNT3)/sqrt(length(QNT3)))) %>%
	arrange(Level, Prefix) %>%
	select(n, mean, lower, upper)
reportDataTable$Order <- NULL
myWeighted <- weighted.var.se(reportDataTable$mean,reportDataTable$n)
reportDataTable$n <- NULL
reportDataTable <- rbind(reportDataTable,myWeighted)
nullHeadings <- c(NA,NA,NA)
reportDataTable <- rbind(nullHeadings,reportDataTable)

reportTextTable <- reportData %>%
	group_by(Order) %>%
	summarise(Semester = Semester[1], Prefix = Prefix[1], Level = Level[1], Outcome = "QNT3", N = length(QNT3), Mean = format(round(mean(QNT3),digits=2),nsmall=2)) %>%
	arrange(Level, Prefix)
reportTextTable$Order <- NULL
reportTextTable$Semester <- as.character(reportTextTable$Semester)
reportTextTable$Prefix <- as.character(reportTextTable$Prefix)
reportTextTable$Level <- as.character(reportTextTable$Level)
headings <- c("Semester","Prefix","Level","Outcome","N","Mean")
reportTextTable <- rbind(headings,reportTextTable)
reportTextTable$Semester <- "Fall 2017"
theSummary <- c("Weighted average",NA,NA,NA,NA,myWeighted)
reportTextTable <- rbind(reportTextTable,theSummary)
```

# Abstract {#abstract -}
> "Assessment is not a spreadsheet; it's a conversation."
> --- Irmeli Halinen

This report is the first quantitative literacy analysis for the new General Education assessment strategy at Ferris State University. A total of 748 student performances on Ferris Learning Outcome 3 were collected over 21 different math courses. Student scores were converted to rubric scores as described above. The overall average rubric score for all students and semesters was 2.69. The mode and median scores were 4 and 3, respectively. The average was not statistically different from the threshold score for competence (2.6) as evaluated with a one-value, two-tailed t-test (t=1.75, df=747, p=0.081). The effect size for the difference between the average and the threshold was tiny (d=0.06). We can infer from this that the overall average rubric score is not practically different than the threshold score. A total of 31 courses were initially registered to submit data for the semester, so we achieved a 67.7% completion rate. According to Banner records, 1,997 students were enrolled in Quantitative Literacy courses in the fall semester. We captured 37.5% of this population with our census.

# Introduction {#intro -}

Assessment is perhaps best viewed as a scholarly activity that is focused upon programmatic improvement. Such scholarly work should be built upon, and contribute to, the relevant professional literature [@Weimer2015]. To emphasize that reality, this report is formatted in the form of a journal article. This report, and ones like it, will be authored, published, and cited in future work to support the development and improvement of the General Education program at Ferris. In addition, reports like this will be the focus of presentations to faculty to suppport continuous improvement of student learning. This report is an analysis of course-level assessment data from 100- and 200-level Math courses in the Fall of 2017. Twenty-one different embedded course assignments were used to evaluate our students' abilities with regard to quantitative literacy outcome QNT3: Apply approaches – Students apply quantitative approaches within contexts to solve problems and draw plausible conclusions.  

# Methods {#methods -}

## Collection of assessment data {-}
Student performances on embedded assignments in 100- and 200-level mathematics courses were analyzed. The content assessed probably varied considerabley. However, all assignments presumably aligned with the QNT3 learning outcome. Individual student scores were collected using the new General Education Natural Sciences "scores" data workbook . Student scores were automatically converted to a rubric score by the workbook using the equivalencies shown in Table \@ref(tab:convert).

```{r convert, echo=FALSE, out.width=4}
Correct <- c("0.0 to 49.0%","50.0 to 59.9%","60.0 to 69.9%","70.0 to 84.9%","85.0 to 100.0%")
Rubric <- c(0,1,2,3,4)
Interpretation <- c("Unsatisfactory","Beginning","Developing","Proficient","Advanced")
tableData <- as.data.frame(cbind(Correct,Rubric,Interpretation))
colnames(tableData) <- c("Percent Correct","Rubric","Interpretation")
knitr::kable(tableData, caption="Conversion of percentages to rubric scores", align=c("l","c","r"), booktabs=TRUE)

```

These workbook files contain personally identifiable information (PII) and are, therefore, subject to FERPA regulations. For this reason, they are not directly shared. Instead, they are permanently housed within the 2017-08 folder under Core Competency: Natural Sciences in TracDat.

## De-identification of student data {-}
Copies of the `r max(reportData$Order)` data files were downloaded from TracDat. An R aggregator script was used to read the data from these data sheets and concatenate it into one data set in a destructive process -- the downloaded copies were deleted in the process. Student names and identification numbers were redacted and each student's entry was given a unique eight-digit identifier - the Record.Key. These keys may be used for longitudinal studies in the future. The algorithm used is kept in an encrypted site and shared with _no one_. The de-identified data set contains `r length(reportData$Order)` student entries and is formatted as a comma-delimited text file (reportData.csv).

## Data provenance {-}
Data provenance refers to a system that permits tracking of the origin, movement, modification, and utilization of data sets [@Buneman2001]. The provenance of General Education data will be explicitly declared to facilitate the reproducibility and extensibility of these studies.

### Location of public website files {-}
All files related to this report can be found online at the Open Science Framework [@Nosek2012]. This site contains all of the files needed to reproduce this report from the de-identified data set. The site's url is [https://osf.io/t6u8m/](https://osf.io/t6u8m/).

### Session information {-}
This report was written using RStudio [@Rstudio] and the R statistical programming language [@R]. These products are free to download for PC, Macintosh, and Linux operating systems. The following information pertains to the session parameters used to generate this report. If you have trouble reproducing this report, it may be due to different session parameters. You may contact [Dr. Franklund](mailto:CliftonFranklund@ferris.edu) if you need assistance.

```{R Session, echo=FALSE, comment=FALSE, error=FALSE, results='asis'}
pander(sessionInfo())
```

### Processing instructions {-}
This project produced a computationally reproducible assessment report (this document). Anyone wishing to recreate this report from the source document will need to install the following on their computer:

1. [An installation of the R programming language](https://www.r-project.org)
2. [An installation of the RStudio IDE](https://www.rstudio.com/products/rstudio/download3/)
3. [An installation of LaTeX](https://www.latex-project.org/get/)

The necessary source files include the de-identified data set (BIOL200Data.csv), Rmarkdown code files (index.Rmd, 01-Introduction.Rmd, 02-Methods.Rmd, 03-Results.Rmd, 04-Discussion.Rmd, and 05-References.Rmd), bibtex reference file (references.bib), and custom art file in the /art folder.

To process the files, you must first open the project in RStudio. Click on the "Build Book" button in the Build menu. Bookdown allows you to build this project as git_book (html site), pdf_book (via LaTeX), or epub_book (compatible with iBooks and other e-book readers).

### Citation of this work {-}
All of the de-identified data, analysis code, and documentation that constitute this report project may be freely used, modified, and shared. The de-identified data set, BIOL200Data.csv, is released under the Creative Commons [CC0 license](https://creativecommons.org/publicdomain/zero/1.0/). All documentation, including README.md, Codebook.md, and this report, are released under the Creative Commons [CC-BY](https://creativecommons.org/licenses/by/4.0/) licence. Any questions, comments, or suggestions may be sent to [Dr. Franklund](mailto:CliftonFranklund@ferris.edu).

# Results {#results -}

This document itself is the primary result of the project. It will be shared with members of the General Education Committee, Academic Senate, and the Department of Mathematics at Ferris State University. Their comments and suggestions will be included in the Discussion.

```{r ttest, echo=FALSE, message=FALSE, comment=NA, results='asis'}
scoreResults <- t.test(reportData$QNT3, mu=2.6)
```

## Summary statistics {-}
A total of `r length(reportData$QNT3)` student performances on Ferris Learning Outcome 3 were collected over `r max(reportData$Order)` different math courses. Student scores were converted to rubric scores as described above. The overall average rubric score for all students and semesters was `r round(mean(reportData$QNT3),2)`. The mode and median scores were `r round(myMode(reportData$QNT3)[1],2)` and `r round(median(reportData$QNT3),2)`, respectively. The average was not statistically different from the threshold score for competence (2.6) as evaluated with a one-value, two-tailed t-test (t=`r round(scoreResults$statistic,2)`, df=`r round(scoreResults$parameter,2)`, p=`r format(scoreResults$p.value, digits=2)`). The effect size for the difference between the average and the threshold was tiny (d=`r round(scoreResults$statistic/sqrt(scoreResults$parameter),2)`). We can infer from this that the overall average rubric score is not practically different than the threshold score. A total of 31 courses were initially registered to submit data for the semester, so we achieved a 67.7% completion rate. According to Banner records, 1,997 students were enrolled in Quantitative Literacy courses in the fall semester. We captured 37.5% of this population with our census.

```{block, type='question'}
Do you think that this level of performance on the outcome is sufficient? Or, do you think that there is a need to increase this level?
```


```{r histogram, echo=FALSE, message=FALSE, results='hide', fig.width=6, fig.align='center', fig.cap="A histogram of the distribution of individual rubric score frequencies over all 21 reports."}
distribution <- table(reportData$QNT3)
barplot(distribution, ylim=c(0,300), las=1, xlab="", ylab="Overall Frequency", axis.lty = 1, col="firebrick", cex.axis = 0.85, cex.lab = 0.85, cex.names=0.85)
mtext(side = 1, text = "Rubric Score on QNT3", line = 1.8, cex=0.85)
```

The distribution of all rubric scores is shown in Figure \@ref(fig:histogram). This distribution exhibited a moderate negative skew (skew = `r round(skewness(reportData$QNT3, na.rm = TRUE),2)`). This result may simply indicate that the teaching, materials, and student learning are all functioning well when the scores are viewed in aggregate. A total of `r sum(reportData$QNT3 >= 3)` students (`r round(sum(reportData$QNT3 >= 3)/length(reportData$QNT3)*100,1)`%) met or exceeded the competence threshold over the semesters investigated. 

```{block, type='question'}
Do you think that this distribution of rubric scores seems right? Are there too many ones and fours, or is this distribution expected in these courses?
```

The distribution of rubric scores by course is shown in Figure \@ref(fig:barplot). There are rather obvious differences in both the distribution of rubric scores and class sizes between semesters. 

```{r barplot, echo=FALSE, message=FALSE, fig.cap="A barplot showing the distribution of rubric scores broken down by course.", fig.width=6, fig.align='center'}
bySemester <- as.matrix(table(reportData$QNT3,reportData$Order))
bySemester <- bySemester[,ncol(bySemester):1]
semesterTotals <- apply(bySemester, 2, sum)
for(dummy in 1:ncol(bySemester)){
	bySemester[,dummy] <- bySemester[,dummy]/semesterTotals[dummy]*100
}
myLabels <- c(rep("Math 100",18),rep("Math 200",3))
#col <- c("firebrick","red","yellow","aquamarine","darkgreen")
col <- c("#a50f15","#de2d26","#fb6a4a","#fcae91","#fee5d9")

par(mar=c(4,8,3,2)+0.1)
barplot(as.matrix(bySemester),
	col=col,
	horiz = TRUE,
	xlab="",
	ylab="",
	names.arg=myLabels,
	xlim=c(0,100),
	las=1,
	cex.axis = 0.75,
	cex.lab=0.75,
	cex.names = 0.75)
mtext("Relative Frequency of Rubric Scores",side=1,line=2, cex = 0.75)
mtext("Reporting Courses",side=2,line=5, cex = 0.75)
add_legend("top",                             # Add a legend to the plot
       legend=c("0","1","2","3","4"),          # Text for the legend
       fill=col,                               # Fill for boxes of the legend
       title="Rubric Score",
       bty="n",
       cex = 0.75,
       horiz = TRUE)                           # Fill for boxes of the legend
```

```{block, type='question'}
What do you think accounts for the variability in scoring from course to course. Is this a problem in measurement, or does the variation reflect real differences in the course populations?
```

## Meta-analysis {-}
Meta-analysis of the student performance was performed using R [@TQMP11-1-37]. This analysis resulted in a weighted average of rubric scores. This value was calculated using formula \@ref(eq:weightX). The value $X_{i}$ average rubric scores for the semesters, while $P_{i}$ is the weighting factor (student enrollment).

\begin{equation}
\bar{X}_w = \frac{\sum X_i P_i}{\sum P_i}
(\#eq:weightX)
\end{equation}

The confidence interval for the weighted mean was calculated using the weighted variance. However, the weighted variance is actually not simple to calculate. Several different methods have been compared to bootstrapping [@Gatz1995a]. The most accurate method was initially described by Cochran [@Cochran1977] and that one was used in this study. The calculation to obtain the weighted variance is shown in formula \@ref(eq:weightV).

\begin{equation}
\begin{split}
(SEM_w)^2 = \frac{n}{(n-1)(\sum P_i)^2}\big[ \sum(P_i X_i - \bar{P}\bar{X}_w)^2 \\
- 2\bar{X}_w \sum(P_i - \bar{P})(P_i X_ i - \bar{P} \bar{X}_w) + \bar{X}_w^2 \sum(P_i - \bar{P})^2 \big]
\end{split}
(\#eq:weightV)
\end{equation}


```{r forest, echo=FALSE, message=FALSE, warning=FALSE, results='hide', fig.cap="A forest plot of the average scores for each course with a weighted mean estimate for the entire period investigated. Error bars indicate the 95% confidence intervals."}

align <- c("c","c","c","c","c","c")
forestplot(reportTextTable, reportDataTable,
	   new_page = FALSE,                             # Image on one page
	   is.summary=c(TRUE,rep(FALSE,21),TRUE),        # Bold for heading and summary lines
	   boxsize = .3,                                 # Set symbol size
	   xlog=FALSE,                                   # Linear scale
	   xticks = c(0,1,1.8,2.6,3.4,4),                        # Ticks at the rubric values
	   zero = 2.6,                                   # Set threshold value
	   grid = gpar(lty=3, col="#333333", lwd=1.25),  # Make vertical lines gray dots
	   xlab = "\nMean rubric score ± 95% CI",        # Label x-axis
	   #title = "Performance on Scientific Understanding Outcome #1 Based Upon Lecture Exam 1",
	   align = align,                                # Center all text columns in table
	   colgap = unit(1, 'mm'),                       # Tighten up the columns
	   graphwidth = unit(70, 'mm'),                  # Make the plot 80mm wide
	   graph.pos=ncol(reportTextTable),                    # Move average values after the plot
	   hrzl_lines = list("1" = gpar(lty=1),          # Add horizontal lines
		"2" = gpar(lty=1),
	   	"20" = gpar(lty=1),
	   	"23" = gpar(lty=1)),
	   txt_gp = fpTxtGp(label=gpar(cex=.75), xlab = gpar(cex=0.75), ticks = gpar(cex=0.75)),
	   col=fpColors(box="firebrick",line="black", summary="firebrick", zero="gray50"))


```

```{r weightedT, echo=FALSE, message=FALSE, warning=FALSE, results='asis'}
averages <- reportData %>%
	group_by(Order) %>%
	summarise(n = length(QNT3), mean = mean(QNT3))

weightedT <- wtd.t.test(averages$mean, 2.6, averages$n)
```

A forest plot of the meta-analysis is shown in Figure \@ref(fig:forest). In this representation, each semester is illustrated as a separate line. The mean and 95% confidence intervals for each semester are plotted in the right panel and their associated meta-data are given in the table to the left. The weighted average of all the data is plotted at the bottom of the figure. The width of the diamond indicates the 95% confidence interval.

```{block, type='question'}
Do you think that this reflects the actual status of student performance in these courses? 
```

The rubric scale can be conceptually divided into five areas as shown in Table \@ref(tab:regions). Of the `r max(reportData$Order)` semesters, `r sum(averages$mean >= 2.6)` fell in the proficient range, `r sum(averages$mean >= 1.8 & averages$mean < 2.6)` fell in the developing range, and `r sum(averages$mean < 1.8)` fell in the beginning range. The weighted mean score was not significantly different from the threshold of competence as judged by a weighted, one-factor, two-tailed t-test (t=`r round(weightedT$coefficients[1],2)`, df=`r weightedT$coefficients[2]`, p=`r round(weightedT$coefficients[3],2)`). We can conclude that the weighted average score is practically equivalent to the competency threshold score.

```{r regions, echo=FALSE, out.width=4}
Average <- c("0.00 to 0.99","1.00 to 1.79","1.80 to 2.59","2.60 to 3.39","3.40 to 4.00")
Interpretation <- c("Unsatisfactory","Beginning","Developing","Proficient","Advanced")
newTable <- as.data.frame(cbind(Average,Interpretation))
names(newTable) <- c("Average Score","Interpretation")
knitr::kable(newTable, caption="Interpretation of average rubric scores", align=c("l","r"), booktabs=TRUE)
```

## Performance by Course Level {-}
```{r level, echo=FALSE, message=FALSE, warning=FALSE, results='hide', fig.cap="A comparison of student performance on QNT3 by course level. Error bars indicate the 95% confidence intervals."}
lev_sum <- summarySE(reportData, measurevar="QNT3", groupvars=c("Level"))

ggplot(data = lev_sum, aes(x = Level, y = QNT3)) +
	geom_bar(stat = "summary", fun.y = "mean", fill="firebrick") +
	ylim(0, 4) +
	geom_errorbar(aes(ymin=QNT3-ci, ymax=QNT3+ci), width=.2) +
	geom_text(aes(label=paste("n =",N)), vjust=10, color="white", size=4) +
	theme_light()
```

```{r ttest2, echo=FALSE, message=FALSE, comment=NA, results='asis'}
levelResults <- t.test(reportData$QNT3 ~ reportData$Level)
```

The average rubric scores for 100-level and 200-level courses was `r round(mean(lev_sum$QNT3[1]),2)` and `r round(mean(lev_sum$QNT3[2]),2)`, respectively . Both of these averages are near the threshold value of 2.6. The difference between these scores was not statistically different as evaluated with a two-tailed t-test (t=`r round(levelResults$statistic,2)`, df=`r round(levelResults$parameter,2)`, p=`r format(levelResults$p.value, digits=2)`). The effect size for the difference between was tiny (d=`r round(levelResults$statistic/sqrt(levelResults$parameter),2)`). We can infer from this that there was no evidence of a measurable difference in student performance by course level.

```{block, type='question'}
The performance in 200-level students is no better than that of 100-level students. Is this a concern, or are these results expected for these courses? Why or why not?
```

## Performance by Standard Measure {-}
```{r measure, echo=FALSE, message=FALSE, warning=FALSE, results='hide', fig.cap="A comparison of student performance on QNT3 by standard meeasure used. Error bars indicate the 95% confidence intervals."}
measure_sum <- summarySE(reportData, measurevar="QNT3", groupvars=c("Standard.Measure"))

ggplot(data = measure_sum, aes(x = Standard.Measure, y = QNT3)) +
	geom_bar(stat = "summary", fun.y = "mean", fill="firebrick") +
	ylim(0, 4) +
	geom_errorbar(aes(ymin=QNT3-ci, ymax=QNT3+ci), width=.2) +
	geom_text(aes(label=paste("n =",N)), vjust=10, color="white", size=4) +
	theme_light()
```

A one-way ANOVA was used to compare the rubric scores by student gender (Table \@ref(tab:anovameasure)). There was not a statistically significant difference between the student scores. So, there is currently no evidence to suggest that there is a difference between students of different standard measures on the Quantitative Literacy learning outcome.

```{r anovameasure, echo=FALSE, results='asis'}
myMeasureModel <- lm(QNT3 ~ Standard.Measure, data = reportData)

knitr::kable(anova(myMeasureModel),caption="One-way ANOVA analysis of scores by standard measure used")
```

```{block, type='question'}
Do you think that the standard measures used in this cycle are the best ones for the outcome being addressed? Are the results surprising to you?
```

## Performance by Student Gender {-}
```{r gender, echo=FALSE, message=FALSE, warning=FALSE, results='hide', fig.cap="A comparison of student performance on QNT3 by student gender. Error bars indicate the 95% confidence intervals."}
gen_sum <- summarySE(reportDataGender, measurevar="QNT3", groupvars=c("Gender"))

ggplot(data = gen_sum, aes(x = Gender, y = QNT3)) +
	geom_bar(stat = "summary", fun.y = "mean", fill="firebrick") +
	ylim(0, 4) +
	geom_errorbar(aes(ymin=QNT3-ci, ymax=QNT3+ci), width=.2) +
	geom_text(aes(label=paste("n =",N)), vjust=10, color="white", size=4) +
	theme_light()
```

A one-way ANOVA was used to compare the rubric scores by student gender (Table \@ref(tab:anovagen)). There was not a statistically significant difference between the student scores. So, there is currently no evidence to suggest that there is a difference between students of different genders on the Quantitative Literacy learning outcome.

```{r anovagen, echo=FALSE, results='asis'}
myGenModel <- lm(QNT3 ~ Gender, data = reportData)

knitr::kable(anova(myGenModel),caption="One-way ANOVA analysis of scores by self-identified student gender")
```

```{block, type='question'}
Do these results seem right to you? Are we reaching all gender groups as well as we ought? Or, do you think that these scores are an anomaly due sampling bias?
```

## Performance by Student Race {-}
```{r race, echo=FALSE, message=FALSE, warning=FALSE, results='hide', fig.cap="A comparison of student performance on QNT3 by student race. Error bars indicate the 95% confidence intervals."}

race_sum <- summarySE(reportDataGender, measurevar="QNT3", groupvars=c("`Race/Ethnicity`"))

ggplot(data = race_sum, aes(x = `Race/Ethnicity`, y = QNT3)) +
	geom_bar(stat = "summary", fun.y = "mean", fill="firebrick") +
	ylim(0, 4) +
	geom_errorbar(aes(ymin=QNT3-ci, ymax=QNT3+ci), width=.2) +
	geom_text(aes(label=paste("n =",N)), vjust=10, color="white", size=4) +
	theme_light()

```

A one-way ANOVA was used to compare the rubric scores by student race (Table \@ref(tab:anova)). There was not a statistically significant difference between the student scores. So, there is currently no evidence to suggest that there is a difference between students of different races on the Quantitative Literacy learning outcome. Note however, that the sample size for most races other than "white" are rather small yet.

```{r anova, echo=FALSE, results='asis'}
myModel <- lm(QNT3 ~ `Race/Ethnicity`, data = reportData)

knitr::kable(anova(myModel),caption="One-way ANOVA analysis of scores by student race")
```

```{block, type='question'}
Do these results seem right to you? Are we reaching all minority groups as well as we ought? Or, do you think that these scores are an anomaly due to the small sample sizes?
```


## Performance by PELL Eligibility {-}
```{r pell, echo=FALSE, message=FALSE, warning=FALSE, results='hide', fig.cap="A comparison of student performance on QNT3 by student PELL eligibility. Error bars indicate the 95% confidence intervals."}

pell_sum <- summarySE(reportData, measurevar="QNT3", groupvars="PELL")

ggplot(data = pell_sum, aes(x = PELL, y = QNT3)) +
	geom_bar(stat = "summary", fun.y = "mean", fill="firebrick") +
	ylim(0, 4) +
	geom_errorbar(aes(ymin=QNT3-ci, ymax=QNT3+ci), width=.2) +
	geom_text(aes(label=paste("n =",N)), vjust=10, color="white", size=4) +
	theme_light()

```

```{r ttest3, echo=FALSE, message=FALSE, comment=NA, results='asis'}
pellResults <- t.test(reportData$QNT3 ~ reportData$PELL)
```

The average rubric scores for Pell-eligible and non-Pell-eligible students was `r round(mean(pell_sum$QNT3[2]),2)` and `r round(mean(pell_sum$QNT3[1]),2)`, respectively. Both of these averages are slightly above the threshold value of 2.6. The difference between these scores was not statistically different as evaluated with a two-tailed t-test (t=`r round(pellResults$statistic,2)`, df=`r round(pellResults$parameter,2)`, p=`r format(pellResults$p.value, digits=2)`). The effect size for the difference between was tiny (d=`r round(pellResults$statistic/sqrt(pellResults$parameter),2)`). We can infer from this that there was no evidence of a measurable difference in student performance by Pell eligibility.

```{block, type='question'}
The performance of Pell-eligible students is no worse than that of those that are not. Is this a surprise, or are these results expected for these courses? Why or why not?
```

# Discussion {#discussion -}

A novel approach for the collection, aggregation, analysis, and reporting of General Education assessment data has been developed. Computationally reproducible reports can easily be generated and distributed to improve the program over time. A meta-analysis of data collected for the first time from a selection of Mathematics courses. 
Over a total of `r max(reportData$Order)` courses, `r round(sum(averages$mean >= 2.6)/length(averages$mean)*100,1)`% had mean scores considered to be proficient. Of all students in all semesters, `r round(sum(reportData$QNT3 >= 3)/length(reportData$QNT3)*100,1)`% met or exceeded the competence threshold. From these data it is inferred that the students are about at the threshold of competence. Disaggregation of scores by standard measure employed, course level, student race, student gender, and Pell eligibility all failed to find any significant differences in student performance.

## Faculty feedback {-}
This report has been distributed to members of the General Education Committee, Academic Senate, and the Department of Mathematics at Ferris State Univerity.  These individuals were asked to provide their comments, suggestions, and concerns about this report and the processes involved in its creation. Their reflections are captured in the Disqus comments within this report. Summary of their ideas and action steps will eventually be included in an Addendum section to this report.

## Plan of action {-}
After analyzing the data and considering the comments provided in the faculty feedback, the relevant General Education sub-committee members will make one or more recommendations for future work. Some of the possible actions could include:

* No modifications -- continue to gather data
* Convene a training session to get better inter-course reliability
* Suggest modifications to the types of assignments that are used
* Suggest modifications to which data workbooks are used
* Suggest that instructors consider modifying the scope or sequence of instruction
* Modify the learning outcomes themselves
* Modify the competency as a whole

## Acknowledgments {-}

This report was built using Rmarkdown and the bookdown R package. The valuable contributions made by the members of the General Education Committee, Academic Senate, and Department of Biological Sciences are also greatly appreciated.

